COVID Shutdown Regulations on Restaurants Decreases xx% of Average Total Sales in Ontario

### 2021-02-19 outputs/paper/paper.Rmd
Commit for Discussion

---
title: "COVID Shutdown Regulations on Restaurants Decreases xx% of Average Total Sales in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan"
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Since 2020, the world-wide COVID-19 pandemic has greatly influenced public health and healthcare, resulting in governments to impose regulations on social gatherings to eliminate further outbreaks. In Ontario, the provincewide stay at home order and shutdown restrictions on businesses and facilities has limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The experiment results show that shutdown regulations on restaurants have decresed xxx%....."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
```

@@

1. show current covid situations and policy regulations from world wide perspective to Ontario specific.

2. some context of economics effects on Ontario, preferably specific to covid effect on restaurant

3. however, current data include not only the shutdown effect, but also the covid effect (counterfeit counterfactual), Inferring causality will require that you rule out other potential factors that could have affected the outcome under consideration.(close of restaurants etc.) Does not make sense to compare the restaurant situations before COVID and during COVID.

4. As a result, to give an accurate estimate of the counterfactual, we need to identify the treatment & control group (2 cities) under current COVID outbreak, with similar restaurant operating situations. We selected Peterborough and Brantford.(could leave this part to discussion)
SAME IN 3 WAYS: (1).avg characteristics identical (through )
 (2). treatment should not affect the comparison group either directly or indirectly ( by city distance of 200+ km unlikely to travel to another city for a meal) However, there might be overestimate of the reopening effect on restaurants since people living near Peterborough while still having their regions' restaurants shutdown might choose to go to Peterborough, increasing sales of Peterborough restaurants.
 (3). Outcome of units in control group should change the same way as outcomes in treatment group (e.g. Provincial subsidy applied to the whole Ontario restaurant business)

Sampling methods and randomizations: Yixin

Data & Survey:
Discussioin Part:
1.Summarize, then
2. mention the selection of cities (limitations)
3. sampling methods (limitations)
4. Need to address internal validity and external validity
Internal validity:Internal validity means that the estimated impact of the program is
net of all other potential confounding factors—or, in other words, that
the comparison group provides an accurate estimate of the counterfactual,
so that we are estimating the true impact of the program.
Basically true since the selection of these two cities ensures the similarity in turn of either city size , population, etc. and restaurant operating situation (through ).

External validity:
External validity means that the evaluation sample accurately represents
the population of eligible units.

Somehow limited, results may not be extrapolated to the entire population of interest, i.e. Restaurant all over Ontario. This is a single-tier relative small size city level experiment. May not generalized to townships, or cosmopolitan city like Toronto...

5. Tradeoff between reopening (economy) and covid cases (public health and healthcare). Should carefully consider....

# Data
```{r}
citation("cansim")
```

&nbsp; &nbsp; Our data explores shutdown effects on restaurant businesses in Ontario. We analyzed it using R [@citeR], and packages `tidyverse` [@cite_tidyverse], `scales` [@cite_scales], `here` [@cite_here]. We used R packages `cansim` [@cite_cansim] and packagesxxx to decide our intervention and sampling method, and packages `bookdown` [@cite_bookdown], `kableExtra`[@cite_kableextra] to format the document.
## Intervention

@@

&nbsp; &nbsp; However, due to current COVID pandemic, any loosening of shutdown restrictions increases chances of COVID exposures, resulting in challenges on public health and healthcare of reopened regions. It is highly risky and hasty to reopen a large area of Ontario regions for the sole purpose of our experiment. Therefore, we mindfully chose Peterborough, a relative small size Single-tier^[Ontario's Municipal Act, 2001 defines a single-tier municipality as "a municipality, other than an upper-tier municipality, that does not form part of an upper-tier municipality for municipal purposes"] city 125 kilometers northeast of Toronto, as the only region to reopen in Ontario and then compared its restaurant operating situations with those in Brantford, another Single-tier city 105 kilometers southwest of Toronto with similar population. Since demographics, land areas and more importantly, restaurant operating situations between these two cities are quite similar with each other^[Approach to examine the similarities of these two cities are further introduced in the discussion section], we consider these two cities as valid comparison cities that could yield an accurate estimate of restaurant businesses shutdowns. In our experiment, we assigned restaurants in Peterborough as the treatment group and those in Brantford as the control group. We also ensured restaurants within these two cities to have equal opportunities receiving government support (e.g. subsidies, grants, etc.) to estimate the true impact of shutdown effects on restaurant businesses after three-month reopening in Peterborough.

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

# Results

@@

## First discussion point

>>>>

---
title: "COVID Shutdown Regulations on Restaurants Decreases xx% of Average Total Sales in Ontario"
author: "Hong Shi, Hong Pan, Yixin Guan, Babak Mokri"
thanks: "Code and data are available at: https://github.com/honn-ishinn/restaurants_covid_shutdown"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Since 2020, the world-wide COVID-19 pandemic has greatly influenced public health and healthcare, resulting in governments to impose regulations on social gatherings to eliminate further outbreaks. In Ontario, the provincewide stay at home order and shutdown restrictions on businesses and facilities has limited economic activities into a recession. This paper examines the shutdown effect on restaurants by reopening a city in Ontario (Peterborough) and comparing its restaurant operating situation with that of another city (Brantford) after three months reopening. The experiment results show that shutdown regulations on restaurants have decresed xxx%....."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

library(tidyverse)
library(bookdown)
library(palmerpenguins)
library(stringr)
library(ggrepel)
library(here)
library(kableExtra)
```

@@

# Data

&nbsp; &nbsp; Our data explores shutdown effects on restaurant businesses in Ontario. We analyzed it using R [@citeR], and packages `tidyverse` [@cite_tidyverse], `here` [@cite_here]. We used R packages `cansim` [@cite_cansim] and packagesxxx to decide our intervention and sampling method, and packages `bookdown` [@cite_bookdown], `kableExtra`[@cite_kableextra] to format the document.

## Intervention

@@

&nbsp; &nbsp; However, due to current COVID pandemic, any loosening of shutdown restrictions increases chances of COVID exposures, resulting in challenges on public health and healthcare of reopened regions. It is highly risky and hasty to reopen a large area of Ontario regions for the sole purpose of our experiment. Therefore, we mindfully chose Peterborough, a relative small size Single-tier^[Ontario's Municipal Act, 2001 defines a single-tier municipality as "a municipality, other than an upper-tier municipality, that does not form part of an upper-tier municipality for municipal purposes"] city 125 kilometers northeast of Toronto, as the only region to reopen in Ontario and then compared its restaurant operating situations with those in Brantford, another Single-tier city 105 kilometers southwest of Toronto with similar population. Since demographics, land areas and more importantly, restaurant operating situations between these two cities are quite similar with each other^[Approach to examine the similarities of these two cities are further introduced in the discussion section], we consider these two cities as valid comparison cities that could yield an accurate estimate of restaurant businesses shutdowns[@cite_impact]. In our experiment, we assigned restaurants in Peterborough as the treatment group and those in Brantford as the control group. We also ensured restaurants within these two cities to have equal opportunities receiving government support (e.g. subsidies, grants, etc.) to estimate the true impact of shutdown effects on restaurant businesses after three-month reopening in Peterborough.

@@

## Determine the Treatment Group and Control Group

To obtain an accurate estimate of the shutdown effect on restaurant businesses, average characteristics of restaurants in the reopening city and those in the shutdown city need to be identical in the absence of the reopen intervention under current COVID outbreak. Since businesses in the food service industry are highly labor intensive [@cite_labor], the `total employment` on food service between selected cities is preferred to be equal as the identical characteristic. Besides, `population density` ( `Total Population`/ `Total Land Area in squared kilometers`) also has significant organizational life-cycle effects[@cite_density]. At different population density levels, competitions for food services varies, leading to different restaurants mortality rates even without the reopen intervention. As a result, total employment on food service and population density serve as primary restaurant operation characteristics for the selection of treatment city and control city.

```{r selectsetup, include=FALSE}

# Employment data
cleaned_statscan_employment <- readr::read_csv(here::here("inputs/data/cleaned_statscan_employment.csv"))
# Census data
cleaned_2016_census <- readr::read_csv(here::here("inputs/data/cleaned_2016_census.csv"))
```

```{r selectclean, include=FALSE}
select_employment <- 
  cleaned_statscan_employment %>% 
  select(GEO, VALUE) %>% 
  filter(GEO != "Ottawa-Gatineau, Ontario/Quebec" & GEO != "Ottawa-Gatineau, Quebec part, Ontario/Quebec")
# Remove the Gatineau employment data since it is out of Ontario province

# We would like to get the population density in one tibble for determining the selection of cities
select_census <- 
  cleaned_2016_census %>% 
  select( City = `Geographic name`, Population_Density = `Population density per square kilometre, 2016`)

# split the column name in select_employment dataset to remove "Ontario" province separated by comma

select_employment$City <- str_split_fixed(select_employment$GEO, ",", 2)[,1] #[,1] means to only keeps the city names column
# Drop the GEO column 
# Ref link to drop the column: https://stackoverflow.com/questions/6286313/remove-an-entire-column-from-a-data-frame-in-r
select_employment$GEO <- NULL

# change Ottawa-Gatineau into Ottawa since already removed Gatineau 
# Ref link to reassign value:https://www.edureka.co/community/35430/how-change-the-value-variable-using-programming-data-frame
select_employment$City[select_employment$City == "Ottawa-Gatineau"] <- "Ottawa"

unique(cleaned_statscan_employment$SCALAR_FACTOR)
# Multiply the employment value by 1000 since the scalar factor is in thousands 
select_employment$Employment <- select_employment$VALUE * 1000
# Drop the VALUE column
select_employment$VALUE <- NULL

# join two dataset to create a dataset that contains cities with population density and employment in Accommodation and food services available

popdensity_employment <- merge(select_census, select_employment, by = "City")

# filter out Toronto since there is no comparable city with Toronto
popdensity_employment <- 
  popdensity_employment %>% 
  filter( City !="Toronto")
```

```{r selectplot, fig.align='center', fig.cap="Population Density vs Employment on Accommodation and Food Services in Ontario Cities in 2016", fig.width=8, fig.height=4, echo= FALSE, message=FALSE, warning=FALSE}

# Highlight cities chosen for our experiment
highlight_city <- popdensity_employment %>% 
  filter(City =="Brantford"|City =="Peterborough")

ggplot( data = popdensity_employment, mapping = aes(x = Employment, y = Population_Density)) +
  geom_point( )+
  geom_point(data = highlight_city,
             aes(x = Employment, y = Population_Density),
             color="red",
             size = 2)+
  theme_minimal() + 
  geom_label_repel(aes( label = City)) +
  labs(
    title = "Population Density vs Total Employment on Accommodation and Food Services",
    x = "Employment",
    y = "Population Density"
  )
```

We used the employment by industry data and 2016 Census data available at Statistics Canada to identify cities with similar total employment on food service and population density. The above point graph(Figure \@ref(fig:selectplot)) indicates that Brantford and Peterborough have similar characteristics affecting the restaurant operations. In addition, the coefficient of the graph $$\frac{Population\:Density}{Total \:Employment}=\frac{Total\:Population/Total\:Land\:Area}{Total\:Employment}$$, which could be interpreted as the number of people served by a restaurant employee per square kilometers, helps to identify the restaurant operating situation in designated regions. Similar coefficients between Brantford and Peterborough (0.306 vs 0.308) also suggest close restaurant characteristics. Besides, there is neither significant demographics nor land areas difference between these two cities, so we consider these two cities as a valid comparison group and reopened Peterborough as the treatment group while maintaining Brantford shutdown as the control group.

Extra concerns when determining the treatment group and control group:

* The employment data at Statistics Canada uses the NAICS (North American Industry Classification System) standard and assigns employment services to Accommodation and Food Service category. So there is less employment in food services. We took the convention that the ratio between the accommodation employment and the food service employment between Peterborough and Brantford is the same.

* Since the distance between Brantford and Peterborough is more than 200 kilometers apart, residents in Brantford are unlikely to travel such long distance solely for a dine-in meal. So the treatment of reopening Peterborough would not affect the mealing behavior in Brantford. However, residents in shutdown regions near Peterborough might go for dine-in meals, increasing sales in Peterborough restaurants. The actual shutdown effect on restaurants could be overestimated due to possible increase in sales from nearby shutdown regions' residents.

* The mindful comparison between cities' similarities on employment, population density, etc. ensures the internal validity of our experiment so that we could accurately estimate the true impact of the shutdown restriction on restaurant. However, the shutdown effect evaluation may not be generalized to the entire population of interest, i.e. all restaurants in Ontario such as township regions or cosmopolitan city Toronto.
